Data analysis for employee cluster assessment

### 2023-07-13 04. Analysis/scripts/03-02-data-analysis-employee-cluster-assessing-clusters.Rmd
Update 03-02-data-analysis-employee-cluster-assessing-clusters.Rmd
Adding silhouette and gap statistic draft visualizations

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
tictoc::tic()
if(!("arsenal" %in% installed.packages()[,"Package"]))
{
  install.packages("arsenal")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
library(oesrrr)
library(factoextra)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
library(readxl)
library(flextable)
library(ggthemes)
library(tidymodels)
```

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
tictoc::tic()
if(!("arsenal" %in% installed.packages()[,"Package"]))
{
  install.packages("arsenal")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
library(oesrrr)
library(factoextra)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
library(readxl)
library(flextable)
library(ggthemes)
library(tidymodels)
library(cluster)
```

@@

### Average silhouette

```{r}
avg_sil <- function(k) {
  km.res <- kmeans(df, centers = k, nstart = 25)
  ss <- silhouette(km.res$cluster, dist(df))
  mean(ss[, 3])
}

k.values <- 2:15

# extract avg silhouette for 2-15 clusters
avg_sil_values <- map_dbl(k.values, avg_sil)

step_01 <- tibble(
  k = k.values,
  silh = avg_sil_values
)

step_01 %>% 
  ggplot(aes(x = k.values, y = silh)) +
  geom_point() +
  geom_line()

plot(k.values, avg_sil_values,
       type = "b", pch = 19, frame = FALSE, 
       xlab = "Number of clusters K",
       ylab = "Average Silhouettes")

fviz_nbclust(df, kmeans, method = "silhouette") +
  theme_economist()
```

```{r}
df <- iris %>% select(where(is.numeric)) %>% drop_na()
```

```{r}
d %>% 
  select_cluster_cols(in_scale_or_rank = "scale",
                      in_working_or_observation_window = "working",
                      in_weekday_correlation = "correlation telework") %>% 
  select(where(is.numeric)) %>% 
  fviz_nbclust(., kmeans, method = "silhouette") +
  theme_fivethirtyeight()
```

### Gap statistic

```{r}
gap_stat <- d %>% 
  select_cluster_cols(in_scale_or_rank = "scale",
                      in_working_or_observation_window = "working",
                      in_weekday_correlation = "correlation telework") %>%
  select(where(is.numeric)) %>% 
  clusGap(., FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
```

```{r}
fviz_gap_stat(gap_stat)
```
